\systemEnding: Bootstrapping Domain-Specific Languages for Neurally-Guided Bayesian Program Learning

### 2019-01-02 archive.tex
improvements to experiment section

\begin{figure}
\centering  \begin{tabular}{ccc}
     \toprule
&     Unigram&Bigram \\\midrule
     $\mathcal{L}^{\text{posterior}}$&
     \begin{tabular}{l}
       \emph{Three samples:}\\
       \code{(+ 1 0)}\\
\code{(+ (+ 0 0) (+ 1 0))}\\
\code{(+ 1 1)}
\\
63.0\% right-associative; 37.4\% \code{+0}'s
       \end{tabular}
     &
     \begin{tabular}{l}
       \emph{Three samples:}\\
       \code{0}\\
\code{(+ (+ (+ 0 0) (+ 0 1)) 1)}\\
\code{1}\\
55.8\% right-associative; 31.9\% \code{+0}'s
       \end{tabular}
     \\\\
     $\mathcal{L}^{\text{MAP}}$&
     \begin{tabular}{l}
       \emph{Three samples:}\\
       \code{1}\\
       \code{(+ 1 (+ 1 (+ (+ 1 (+ 1 1)) 1)))}\\
       \code{(+ (+ 1 1) 1)}
\\48.6\% right-associative;
       0.5\% \code{+0}'s       
       \end{tabular}
     &
     \begin{tabular}{l}
       \emph{Three Samples:}\\
\code{(+ 1 (+ 1 (+ 1 (+ 1 (+ 1 1)))))}\\
\code{0}\\
\code{(+ 1 (+ 1 (+ 1 1)))}
       \\
       \textbf{97.9\% right-associative};
       2.5\% \code{+0}'s       
       \end{tabular}
\\    \bottomrule 
  \end{tabular}
  \caption{Agent learns to break symmetries in program space only when using both bigram parameterization and $\mathcal{L}^{\text{MAP}}$ objective, associating addition to the right and avoiding adding zero. \% right-associative calculated by drawing 500 samples from $Q$. $\mathcal{L}^{\text{MAP}}$/Unigram agent incorrectly learns to never generate programs with \code{0}'s, while $\mathcal{L}^{\text{MAP}}$/Bigram agent correctly learns that \code{0} should only be disallowed as an argument of addition. Tasked with building programs from \code{+}, \code{1}, and \code{0}. }\label{symmetry}
  \end{figure}

@@

Inspired by the classic AI `copy task' --- where an agent must look at
an image of a tower made of toy blocks and re-create the
tower~\cite{towerCopy} --- we give \system 112 tower `copy tasks'
(Figure~\ref{towerTasks}).  Here the agent observes both an image of a
tower and the locations of each of its blocks, and must write a
program that plans how a simulated hand would build the tower.
These towers are built from Lego-style blocks
that snap together on a discrete grid.%%  (e.g., we do not model gravity
%% and discretize the coordinates of the 
%% blocks).

The system starts out with four domain-specific primitives:
\code{left} and \code{right}, which take an integer $n$ as input and
move the hand left/right $n$ Lego grid spaces; along with
\code{horizontal} and \code{vertical}, which put down
horizontal/vertical Lego blocks.  As with LOGO graphics we also
include our two standard planning primitives (\code{for} and
\code{get/set}).
Starting with this basis
the system can only solve
12.5\% of the copy tasks.
\begin{figure}
  \begin{tabular}{cc}
    \toprule Example learned DSL components&Samples from DSL\\\midrule
    \begin{tabular}{rl}
      Brickwall&\begin{tabular}{l}
        \includegraphics[width = 5cm]{figures/tower/tower_dsl_bricks.png}
      \end{tabular}\\
      Bridge&\begin{tabular}{l}
        \includegraphics[width = 5cm]{figures/tower/tower_dsl_bridge.png}
      \end{tabular}\\
      Staircase&\begin{tabular}{l}
        \includegraphics[width = 5cm]{figures/tower/tower_dsl_staircase.png}
        \end{tabular}
    \end{tabular}&
    \begin{tabular}{l}
      \includegraphics[width = 5cm]{figures/tower/dreams.png}
      \end{tabular}
    \\\bottomrule 
    \end{tabular}
  \caption{Learned subroutines for building towers out of Lego-style blocks. These subroutines act like parametric options~\cite{options},
    giving higher-level building blocks that the agent can use to plan.}
\end{figure}

\subsubsection{Symbolic Regression: Programs from visual input}\label{regressionSection}
Here, the agent observes points along the curve of a function, and
must write a program that fits those points.  We initially equip our
learner with addition, multiplication, and division, and task it with
solving 200 symbolic regression problems, each either a polynomial or
rational function.  The recognition model is a convnet that observes
an image of the target function's graph (Fig.~\ref{visualSpecs},
rightmost) --- visually, different kinds of polynomials and rational
functions produce different kinds of graphs, and so the convnet can
look at a graph and predict what kind of function best explains it.  A
key difficulty, however, is that these problems are best solved with
programs containing real numbers.  Our solution to this difficulty is
to enumerate programs with real-valued parameters, and then fit those
parameters by automatically differentiating through the programs the
system writes and use gradient descent to fit the parameters.  We
define the likelihood model, $\probability[x|p]$, by assuming a
Gaussian noise model for the input/output examples, and penalize the
use of real-valued parameters using the
BIC~\cite{Bishop:2006:PRM:1162264}.

@@

h
\subsection{Quantitative Results on Held-Out Tasks}\label{quantitative}
We evaluate  on held-out testing tasks,
measuring how many
tasks are solved and how long it takes to solve them (Fig.~\ref{learningCurves}).
Prior to any learning,
the system cannot find solutions for most of the tasks,
and those it does solve take a long time;
with more wake/sleep iterations,
we converge upon DSLs and recognition models 
more closely matching the domain.

>>>>

\begin{figure}
  \begin{minipage}[c]{0.3\textwidth}
    \begin{tikzpicture}%[every node/.style={inner sep=0,outer sep=0}]
    \node(l)[draw] at (0,0) {\texttt{$\lambda$ (a)}};
    \node(k)[draw] at ([yshift=-50]l.south) {\texttt{+}};
    \node(o)[draw] at ([xshift=-50,yshift=-50]k.south) {\texttt{9}};
    \node(m)[draw] at ([xshift=50,yshift=-50]k.south) {\texttt{*}};
    \node(x1)[draw] at ([xshift=50,yshift=-50]m.south) {\texttt{a}};
    \node(x2)[draw] at ([xshift=-50,yshift=-50]m.south) {\texttt{a}};

    \draw[->] (l.south)-- node[fill=white,align=center,midway,inner sep=0,outer sep=0]{$Q_{\text{start},\texttt{+},1}(x)$}(k.north);
    \draw[->] (k.south)--node[fill=white,align=center,midway,inner sep=0,outer sep=0]{$Q_{\texttt{+},\texttt{9},1}(x)$}(o.north);
    \draw[->] (k.south)--node[fill=white,align=center,midway,inner sep=0,outer sep=0]{$Q_{\texttt{+},\texttt{*},2}(x)$}(m.north);
    \draw[->] (m.south)--node[fill=white,align=center,midway,inner sep=0,outer sep=0]{$Q_{\texttt{*},\text{var},2}(x)$}(x1.north);
    \draw[->] (m.south)--node[fill=white,align=center,midway,inner sep=0,outer sep=0]{$Q_{\texttt{*},\text{var},1}(x)$}(x2.north);
\end{tikzpicture}
    \end{minipage}\hfill%
  \begin{tabular}{cll}
     \toprule
&     \multicolumn{1}{c}{Unigram}&\multicolumn{1}{c}{Bigram} \\\midrule
     $\mathcal{L}^{\text{posterior}}$&
     \begin{tabular}{l}
       \emph{Three samples:}\\
       \code{(+ 1 0)}\\
\code{(+ (+ 0 0)}\\\code{\phantom{(+ }(+ 1 0))}\\
\code{(+ 1 1)}
\\
63.0\% right-associative\\37.4\% \code{+0}'s
       \end{tabular}
     &
     \begin{tabular}{l}
       \emph{Three samples:}\\
       \code{0}\\
\code{(+ (+ (+ 0 0)}\\\code{\phantom{(+ (+} (+ 0 1)) 1)}\\
\code{1}\\
55.8\% right-associative\\31.9\% \code{+0}'s
       \end{tabular}
     \\\\
     $\mathcal{L}^{\text{MAP}}$&
     \begin{tabular}{l}
       \emph{Three samples:}\\
       \code{1}\\
       \code{(+ 1 (+ 1 (+ (+ 1 }\\\code{\phantom{(+ }(+ 1 1)) 1)))}\\
       \code{(+ (+ 1 1) 1)}
\\48.6\% right-associative\\
       0.5\% \code{+0}'s       
       \end{tabular}
     &
     \begin{tabular}{l}
       \emph{Three Samples:}\\
\code{(+ 1 (+ 1 (+ 1}\\\code{\phantom{(+ 1} (+ 1 (+ 1 1)))))}\\
\code{0}\\
\code{(+ 1 (+ 1 (+ 1 1)))}
       \\
       \textbf{97.9\% right-associative}\\
       2.5\% \code{+0}'s       
       \end{tabular}
\\    \bottomrule 
  \end{tabular}
  \caption{\textbf{Left:} Bigram parameterization of distribution over programs predicted by recognition model.
  Here the program (syntax tree shown above) is \texttt{($\lambda$ (a) (+ 9 (* a a )))}.
Each conditional distribution predicted by the recognition model is written $Q_{\text{parent},\text{child},\text{argument index}}(x)$, where $x$ is a task. \textbf{Right:} Agent learns to break symmetries in program space only when using both bigram parameterization and $\mathcal{L}^{\text{MAP}}$ objective, associating addition to the right and avoiding adding zero. \% right-associative calculated by drawing 500 samples from $Q$. $\mathcal{L}^{\text{MAP}}$/Unigram agent incorrectly learns to never generate programs with \code{0}'s, while $\mathcal{L}^{\text{MAP}}$/Bigram agent correctly learns that \code{0} should only be disallowed as an argument of addition. Tasked with building programs from \code{+}, \code{1}, and \code{0}. }\label{symmetry}
\end{figure}

\begin{figure}
\centering  \begin{tabular}{cll}
     \toprule
&     \multicolumn{1}{c}{Unigram}&\multicolumn{1}{c}{Bigram} \\\midrule
     $\mathcal{L}^{\text{posterior}}$&
     \begin{tabular}{l}
       \emph{Three samples:}\\
       \code{(+ 1 0)}\\
\code{(+ (+ 0 0) (+ 1 0))}\\
\code{(+ 1 1)}
\\
63.0\% right-associative; 37.4\% \code{+0}'s
       \end{tabular}
     &
     \begin{tabular}{l}
       \emph{Three samples:}\\
       \code{0}\\
\code{(+ (+ (+ 0 0) (+ 0 1)) 1)}\\
\code{1}\\
55.8\% right-associative; 31.9\% \code{+0}'s
       \end{tabular}
     \\\\
     $\mathcal{L}^{\text{MAP}}$&
     \begin{tabular}{l}
       \emph{Three samples:}\\
       \code{1}\\
       \code{(+ 1 (+ 1 (+ (+ 1 (+ 1 1)) 1)))}\\
       \code{(+ (+ 1 1) 1)}
\\48.6\% right-associative;
       0.5\% \code{+0}'s       
       \end{tabular}
     &
     \begin{tabular}{l}
       \emph{Three Samples:}\\
\code{(+ 1 (+ 1 (+ 1 (+ 1 (+ 1 1)))))}\\
\code{0}\\
\code{(+ 1 (+ 1 (+ 1 1)))}
       \\
       \textbf{97.9\% right-associative};
       2.5\% \code{+0}'s       
       \end{tabular}
\\    \bottomrule 
  \end{tabular}
  \caption{Agent learns to break symmetries in program space only when using both bigram parameterization and $\mathcal{L}^{\text{MAP}}$ objective, associating addition to the right and avoiding adding zero. \% right-associative calculated by drawing 500 samples from $Q$. $\mathcal{L}^{\text{MAP}}$/Unigram agent incorrectly learns to never generate programs with \code{0}'s, while $\mathcal{L}^{\text{MAP}}$/Bigram agent correctly learns that \code{0} should only be disallowed as an argument of addition. Tasked with building programs from \code{+}, \code{1}, and \code{0}. }\label{symmetry}
  \end{figure}

@@

Inspired by the classic AI `copy task' --- where an agent must look at
an image of a tower made of toy blocks and re-create the
tower~\cite{towerCopy} --- we give \system 112 tower `copy tasks'
(Figure~\ref{towerTasks}).  Here the agent observes both an image of a
tower and the locations of each of its blocks, and must write a
program that plans how a simulated hand would build the tower.
These towers are built from Lego-style blocks
that snap together on a discrete grid.%%  (e.g., we do not model gravity
%% and discretize the coordinates of the 
%% blocks).
The system starts with the same control flow primitives as with
LOGO graphics,
and learns parametric `options' for building blocks towers (Figure~\ref{tower}),
inferring concepts like arches, staircases, bridges and brick walls.
\begin{figure}
  \begin{tabular}{cc}
    \toprule Example learned DSL components&Samples from DSL\\\midrule
    \begin{tabular}{rl}
      Brickwall&\begin{tabular}{l}
        \includegraphics[width = 5cm]{figures/tower/tower_dsl_bricks.png}
      \end{tabular}\\
      Bridge&\begin{tabular}{l}
        \includegraphics[width = 5cm]{figures/tower/tower_dsl_bridge.png}
      \end{tabular}\\
      Staircase&\begin{tabular}{l}
        \includegraphics[width = 5cm]{figures/tower/tower_dsl_staircase.png}
        \end{tabular}
    \end{tabular}&
    \begin{tabular}{l}
      \includegraphics[width = 5cm]{figures/tower/dreams.png}
      \end{tabular}
    \\\bottomrule 
    \end{tabular}
  \caption{\textbf{Left}: Three (out of 19) learned DSL components for building towers out of Lego-style blocks. These components act like parametric options~\cite{options},
    giving higher-level building blocks that the agent can use to plan. \textbf{Right}: 16 random samples, or `dreams', from learned DSL.}\label{tower}
\end{figure}

\subsubsection{Symbolic Regression}\label{regressionSection}
Here, the agent observes points along the curve of a function, and
must write a program that fits those points.  We initially equip our
learner with addition, multiplication, and division, and task it with
solving 200 symbolic regression problems, each either a polynomial or
rational function.  The recognition model is a convnet that observes
an image of the target function's graph (Fig.~\ref{visualSpecs},
rightmost) --- visually, different kinds of polynomials and rational
functions produce different kinds of graphs, and so the convnet can
look at a graph and predict what kind of function best explains it.  A
key difficulty, however, is that these problems are best solved with
programs containing real numbers.  Our solution to this difficulty is
to enumerate programs with real-valued parameters, and then fit those
parameters by automatically differentiating through the programs the
system writes and use gradient descent to fit the parameters.  We
define the likelihood model, $\probability[x|p]$, by assuming a
Gaussian noise model for the input/output examples, and penalize the
use of real-valued parameters using the
BIC~\cite{Bishop:2006:PRM:1162264}.

@@

\subsection{Quantitative Results on Held-Out Tasks}\label{quantitative}
We evaluate  on held-out testing tasks,
measuring how many
tasks are solved and how long it takes to solve them (Fig.~\ref{learningCurves}).
Prior to any learning,
the system cannot find solutions for most of the tasks,
and those it does solve take a long time;
with more wake/sleep iterations,
we converge upon DSLs and recognition models 
more closely matching the domain.
